Appendix: Quantitative, Multispecies Monitoring at a Continental Scale

### 2025-02-27 Appendix_plain_v3.tex
Updates from Overleaf

\begin{align} 
	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \omega_{d}(x,y,d) \\
	 G_{s}(x,y,d) \sim \mathrm{Tweedie}(F_s(x,y,d),p ,\theta_s)
\end{align}
where x, y, and d index longitude, latitude, and depth, respectively.   $\kappa$ is scalar intercept. Both $\eta(x,y)$ and $\omega_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\omega_d \sim  \mathrm{MVN}(0,\sum_{\omega_d})$ (see \cite{anderson2022}) where $\eta$ is a depth-invariant spatial field and $\omega_d$ is a spatial field specific to each depth category. In simple terms, each location-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific. $\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\omega_d$ modulates that estimate for different sampling depths. The $\omega_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$.

>>>>

\begin{align} 
	 log(F_s(x,y,d)) = \kappa + \eta(x,y) + \omega_{d}(x,y,d) \\
	 G_{s}(x,y,d) \sim \mathrm{Tweedie}(F_s(x,y,d),p ,\theta_s)
\end{align}
where x, y, and d index longitude, latitude, and depth, respectively.   $\kappa$ is scalar intercept. Both $\eta(x,y)$ and $\omega_d(x,y,d)$ are species- and location-specific realizations of Gaussian Markov random fields, $\eta \sim  \mathrm{MVN}(0,\sum_\eta)$ and $\omega_d \sim  \mathrm{MVN}(0,\sum_{\omega_d})$ (see \cite{anderson2022}) where $\eta$ is a depth-invariant spatial field and $\omega_d$ is a spatial field specific to each depth category. In simple terms, each location-specific estimate of log eDNA concentration is estimated as a sum of the scalar intercept and two spatial fields, one that is common across depths and one that is depth-specific. $\eta$ estimates the smooth, two-dimensional (latitude/longitude) spatial pattern in estimated DNA, while $\omega_d$ modulates that estimate for different sampling depths. The $\omega_d$ fields were allowed to be independent and identically distributed (IID), and therefore there was no forced correlation across depths beyond that provided by the common spatial field $\eta$. Additionally to convert eDNA concentration from from copies/$\mu$L (reaction concentration) to copies/L of water filtered (ecological concentration) we multiply $F_s(x,y,d)$ by 20 which is the conversion value indicating the ratio of filtered volume and the volume of DNA template used in the PCR reaction.
